---
title: "code_chunk_demo"
author: "Jessica Bonin"
date: "10/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_peguin_data, echo = TRUE, eval = TRUE}
require(palmerpenguins)
head(penguins)
```

```{r sample pairplot a}
plot(penguins[, 3:5])
```

```{r my chunck 1}
```

# Show or hide code: is echo = TRUE or FALSE
# Show or hie chunk output in the rendar doc: eval = TRUE or FALSE

```{r sample pairplot b, fig.width=4, fig.asp=1}
plot(penguins[, 3:5])
```

```{r sample pairplot c, fig.width=4, fig.asp=1/1.5}
plot(penguins[, 3:5])
```

# Will only show code, not the plot
```{r sample pairplot d, eval= FALSE}
plot(penguins[, 3:5])
```

# Will only show the figure, not the r code
```{r sample pairplot f, echo= FALSE}
plot(penguins[, 3:5])
```

# Tabset Example {.tabset}

## Tab 1
Tab 1 Contents

## Tab 2
Tab 2 Contents

## Tab 3
Tab 3 Contents

# Next Section
Non-Tabbed Material

# Tabset Pill Example {.tabset .tabset-pills}

## Tab 1
Tab 1 Contents

## Tab 2
Tab 2 Contents

## Tab 3
Tab 3 Contents

# Next Section
Non-Tabbed Material


# Different Penguin Plots {.tabset .tabset-pills}

The following tabs contain different plot types

## Histogram
```{r flipper length histogram, fig.asp=1/1.5}
hist(penguins$flipper_length_mm)
```

## Another Histogram (Bill Length)
```{r bill length histogram, echo = FALSE}
hist(penguins$bill_length_mm)
```

