---
title: 'Final Project: R Reference'
author: "Jessica Bonin"
date: "12/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages
- To load packages, you can install them using **install.packages()**.
- To "turn on" or use a package that you have already installed into r, you can use either **library()** or **require()**.
- The function **here()** lets to tell r where to pull data from your file directory. It starts where your r project is saved, in this case it is in my environmental_data folder.
- To create a data.frame in r, you name your data and tell r where to pull the data from.

```{r}
# Installing a package
# install.packages("palmerpenguins")

# Turning on a package that was previously installed
library(palmerpenguins)
# OR
require(palmerpenguins)

# Creating a data.frame
penguins <- palmerpenguins::penguins

require(here)
ginkgo <- read.csv(here("data", "ginkgo_data_2021.csv"))
```

## Data Structures
### Vectors
You can create numeric, character, or boolean vectors by using **c()**. You can name the values in a vector or name a whole vector by using **names()**.
```{r}
# Creating a vector of numbers
numeric_vector <- c(2, 4, 6, 8)

# Creating a vector of characters
character_vector <- c("who", "do", "we", "appreciate")

# Creating a boolean_vector:
boolean_vector <- c(TRUE, TRUE, FALSE)

# Naming numeric_vector
names(numeric_vector) <- c("two", "four", "six", "eight")
```
### Matrices 
A matrix is a collection of elements of the same data type. You can name rows or columns in a matrix using **rownames()** and **colnames()**.
```{r}
# Matrix with numbers 1-9 in 3 rows
my_matrix <- matrix(1:9, byrow = TRUE, nrow = 3)
# You can also use a vector in place of numbers (instead of 1:9). It can be a defined vector or using c().

# Naming rows and columns
rows <- c("row1", "row2", "row3")
columns <- c("col1", "col2", "col3")
rownames(my_matrix) <- rows
colnames(my_matrix) <- columns 
```
### Length
Get or set the length of an R object using **length()**.
```{r}
length(numeric_vector)
length(boolean_vector)
length(my_matrix)
# Find out how many variables are in a data frame
length(ginkgo) 
length(penguins)
```

### Data Frames
You can create data frames many different ways (see Loading Data and Packages). This is an example of how to create a data frame from scratch (example from **tutorialspoint.com**).
```{r}
emp.data <- data.frame(
  emp_id = c (1:5), 
  emp_name = c("Rick","Dan","Michelle","Ryan","Gary"),
  salary = c(623.3,515.2,611.0,729.0,843.25), 
   
  start_date = as.Date(c("2012-01-01", "2013-09-23", "2014-11-15", "2014-05-11",
      "2015-03-27")),
  stringsAsFactors = FALSE
)
print(emp.data)
```
### Data Frame Counts
Use **nrow()** to find out the number of rows/observations in a data frame. Use **ncol()** to find out the number of columns/variables in a data frame. Use **dim()** to find out the dimensions of a data frame.
```{r}
nrow(ginkgo)
ncol(ginkgo)
dim(ginkgo)
```

## Subsetting
You may want to select elements from a vector, matrix, or data frame. There are a few ways you can do this...
```{r}
# Using $ to select a column of data
ginkgo$seeds_present

# Using [] to subset by position:
# Select first row of the ginkgo data
ginkgo[1,]

# Select the element in row 2, column 3
ginkgo[2,3]

# Select the 3rd column of the ginkgo data
ginkgo[,3]

# Using subset() function to retrieve all the data for Adelie penguins (in the species column) from the peuguins dataset.
Adelie <- subset(penguins, species == "Adelie")
Gentoo <- subset(penguins, species == "Gentoo")
Chinstrap <- subset(penguins, species == "Chinstrap")
```

## Numerical Data Exploration
There are functions in r that you can use for data calculations. Using **na.rm** will excuse missing data to allow the function to run.
```{r}
# Summary tells you a summary of the results (compare it to the mean() function).
summary(penguins$body_mass_g)
mean(penguins$body_mass_g, na.rm = TRUE)
sd(penguins$body_mass_g, na.rm = TRUE)
```

## Graphical Data Exploration
There are many ways to graph data in r. The most common I have used are: scatterplots, histograms, and boxplots.

### Scatterplots
Scatterplots graph data as (x,y) points. They are helpful in evaluating trends in the relationships of the data.
```{r}
# col = color
# pch = symbol
# cex = size
plot(ginkgo$max_width, ginkgo$max_depth, col = "cadetblue4",
     pch = 1, cex = 2, main = "Ginkgo Leaf Max Width by Max Depth", xlab = "Max Width (mm)", ylab = "Max Depth (mm)", xlim = c(0,131), ylim = c(0,150))
```

### Histograms
Histograms graph data as bar charts based on the frequency of data. They are good at evaluating distribution. Use **par()** to put all histograms in one figure.
```{r, fig.height=7, fig.width=7}
par(mfrow=c(2,2))
hist(penguins$flipper_length_mm, 
     main = "Penguins Flipper Length", 
     xlab = "Flipper Length (mm)", col = "goldenrod3",
     breaks = 15, xlim = c(170,235))
hist(Adelie$flipper_length_mm, 
     main = "Adelie Flipper Length", 
     xlab = "Flipper Length (mm)", col = "darkslategray4",
     breaks = 15, xlim = c(170,210))
hist(Gentoo$flipper_length_mm,
     main = "Gentoo Flipper Length", 
     xlab = "Flipper Length (mm)", col = "coral3",
     breaks = 15, xlim = c(200,235))
hist(Chinstrap$flipper_length_mm,
     main = "Chinstrap Flipper Length", 
     xlab = "Flipper Length (mm)", col = "darkseagreen4",
     breaks = 15, xlim = c(170,220))
```

### Boxplots
Boxplots are good for looking at the spread of the data as well as the distribution. They are also helpful when comparing variables of the data set.
```{r, fig.height=7, fig.width=9}
par(mfrow=c(1,2))
# Simple boxplot
boxplot(ginkgo$petiole_length, main = "Ginkgo Petiole Length", 
        ylab = "Petiole Length (mm)", col = "blue")

# Conditional boxplot
boxplot(max_depth ~ seeds_present, data = ginkgo,
        main = "Ginkgo Max Depth \n by Seed Presence", xlab = "Seeds Present", 
        ylab = "Max Depth (mm)", col = c("coral3", "darkseagreen3"))
```

## Disctribution Functions
- The Normal Distribution: Density, distribution function, quantile function and random generation for the normal distribution with mean equal to mean and standard deviation equal to sd.
- The Binomial Distribution: Density, distribution function, quantile function and random generation for the binomial distribution with parameters size and prob. This is conventionally interpreted as the number of ‘successes’ in size trials.
- Some of these examples came from **tutorialspoint.com**. 

### dnorm(): Probability Density
Using **dnorm()** gives height of the probability distribution at each point for a given mean and standard deviation.
dnorm(x, mean = 0, sd = 1, log = FALSE). x = a vector of **numbers**.
```{r}
# Create a sequence of numbers between -10 and 10 incrementing by 0.1.
x <- seq(-10, 10, by = .1)

# Choose the mean as 2.5 and standard deviation as 0.5.
y <- dnorm(x, mean = 2.5, sd = 0.5)
plot(x,y)

# Finding the a single probability
dnorm(0.5, mean = 2.5, sd = 0.5)
```

### pnorm(): Cumulative Distribution Function
Using **pnorm()** gives the probability of a normally distributed random number to be less that the value of a given number.
pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE). q = a vector of **numbers**.
```{r}
# Create a sequence of numbers between -10 and 10 incrementing by 0.2.
x <- seq(-10,10,by = .2)
 
# Choose the mean as 2.5 and standard deviation as 2. 
y <- pnorm(x, mean = 2.5, sd = 2)
plot(x,y)

# Finding the a single probability
pnorm(3, mean = 2.5, sd = 2)
```

### qnorm(): Quantile Function
Using **qnorm()** takes the probability value and gives a number whose cumulative value matches the probability value.
qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE). p = a vector of **probabilities**.
```{r}
# Create a sequence of probability values incrementing by 0.02.
x <- seq(0, 1, by = 0.02)

# Choose the mean as 2 and standard deviation as 3.
y <- qnorm(x, mean = 2, sd = 1)
plot(x,y)

# Finding the a single probability
qnorm(0.43, mean = 2, sd = 1)
```

### dbinom()
Using **dbinom()** gives the probability density distribution at each point.
dbinom(x, size, prob, log = FALSE). x = a vector of **numbers**.
```{r}
# Create a sample of 50 numbers which are incremented by 1.
x <- seq(0, 50, by = 1)

# Create the binomial distribution.
y <- dbinom(x, 50, 0.5)
plot(x,y)

# Probability of observing 5 out of 9 with a 1/3 probability of observation
dbinom(5, 9, 1/3)
```

### pbinom()
Using **pbinom()** gives the cumulative probability of an event. It is a single value representing the probability.
pbinom(q, size, prob, lower.tail = TRUE, log.p = FALSE). q = a vector of **numbers**.
```{r}
# Probability of getting 26 or less heads from a 51 tosses of a coin.
pbinom(26, 51, 0.5)

# Probability of getting 26 or more heads from a 51 tosses of a coin.
1-pbinom(25, 51, 0.5)
```

### qbinom()
Using **qbinom()** takes the probability value and gives a number whose cumulative value matches the probability value.
qbinom(p, size, prob, lower.tail = TRUE, log.p = FALSE). p = a vector of **probabilities**.
```{r}
# How many heads will have a probability of 0.25 will come out when a coin
# is tossed 51 times.
qbinom(0.25, 51, 1/2)
```